---
title: "pineplot: Cleveland heart disease dataset"
author: "Daniel Hogan"
date: "July 20, 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cleveland_heart_disease}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pineplot)
library(gtable)
library(grid)
library(gridExtra)
```

# Loading the Data

For this example we will be using the Cleveland heart disease dataset. The dataset contains 14 attributes (as columns) across 303 subjects (as rows). The attributes were selected for their relevance to heart disease and include variables such as age, sex, cholesteral level, and maximum heart rate achieved.

```{r}
data(heart)
colnames(heart)
attr(heart, "format") # column information
```

# Preprocessing

We will be examining the interaction between several variables while controlling for factors of sex and age. Levels for the sex attribute are already defined (1 = male; 0 = female). For the age attribute we define 3 levels (i.e. age groups) using the `cut` function. The dataframe is then partitioned by sex and age and correlation matrices are calculated for each partition.

```{r}
heart['age'] <- cut(heart[['age']], 3)
heart_mat <- heart[c('trestbps', 'chol', 'fbs', 'restecg', 'thalach')]
heart_subsets <- split(heart_mat, list(heart$sex, heart$age))
heart_subsets <- lapply(heart_subsets, cor)
```

# Annotation Functions

We define a callback function that generates a text grob (GRaphical OBject) given the heatmap objects and a name argument. The callback function will be used to label the heatmaps.

```{r}
add_label <- function(panel, name){
  text <- paste0('Age:\n', sub('.*\\.', '', name))
  textGrob(text, x=0, y=.8, just='left')
}
```

# Pineplot Generation

Two pineplots are generated, one for females and one for males. The pineplots are stored in a list for later arrangment using the grid package. When calling generate_pineplot, we pass the callback function as an argument. The \code{\link{grid.arrange}} function is used to lay the pine plots out side-by-side for easy comparison.

```{r fig.align='center', fig.height=5, fig.width=5, out.width='100%'}
pineplots <- list()
for (sex in c(0, 1)) {
  sex_indices <- as.numeric(sub('\\..*', '', names(heart_subsets)))
  heart_by_sex <- heart_subsets[which(sex_indices == sex)]
  pp <- generate_pineplot(
    heart_by_sex,
    customize_fn = add_label,
    breaks=c(-1, -.5, 0, .5, 1),
    limits=c(-1, 1),
    low="blue", mid="yellow", high="red",
    legend_scale=1.2
  )
  pineplots[[sex+1]] <- pp
}
grid.arrange(pineplots[[1]], pineplots[[2]], nrow=1)
```
